\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{plain}

I am deeply grateful for the patience and support of my supervisors, Andreas and Alexander. Thank you for being super flexible along the way, though this work took significantly longer than planned, and for investing your valuable time in discussions or correcting the thesis. I also want to thank Bernd Kaifler for providing the CORAL temperature measurements and the whole Middle Atmosphere group at the DLR for fruitful discussions and instructive inputs. The same applies to the numerical modeling group of Alexander Gohm at the University of Innsbruck.\\
I am grateful that the IPA, particularly Markus Rapp and Andreas Fix, gave me the opportunity to attend events like the SPARC Symposium in Frankfurt or the Southtrac-GW workshop in Bad Tölz. These have been exciting experiences.

At this point, I also want to thank the team of the avalanche warning service I worked with in between. Though snow at the surface and waves in the stratosphere do not have much in common, the methods we researchers use to analyze them all the more. Not only did I enjoy the time working there, the results and figures of my thesis also benefited from the interlude. Thank you, Machtl, for fun-filled evening coding sessions and for being an amazing wingman in and out of the office. And thank you, Christoph, for trusting us with awsome projects and being a great mentor.

My emotional thanks go to my family. You are my harbour when I make things too complicated from time to time. With you, life becomes easy again, and I am grateful I can always rely on you. And cheers, Valli, for being my favorite distraction while finishing this thesis.

% My emotional thanks go to my family, of course. Even though I rarely give the impression that I need help, I am grateful that I can always rely on you. 

% am very grateful that you gave me the opportunities to attend international conferences like the American and European Geoscience Unions and to do exciting trips to places like Kühlungsborn, Kiruna, S ̃ao José dos Campos or to the end of the world, Río Grande. 

% Thank you for your valuable time that you always in- vested in correcting my work. I assume it has not always been easy. Furthermore, I would like to thank the “Middle Atmosphere” matrix group at the IPA in Oberpfaffenhofen for countless discussions with colleagues such as Thomas Birner, Andreas Dörnbrack, Hella Garny, Sonja Gisinger, Stefanie Knobloch, Tyler Mixa and Henrike Willms. Thank you, Andreas, for taking me under your wings more and more during the course of my doctoral studies

% << Alex >>
% Now it is time to thank all people who have contributed to your work and who
% have supported you during your study. Do not forget to mention all relevant data
% providers and funding agencies (also provide the grant numbers).
